---
title: "Project Part I: Statistical Report"
author: "Nattapon pronsukwitoon 48995940"
format: 
  html: default
  pdf: default
params:
  datafile: "E:/StatGit/project2025.csv"
execute: 
  echo: false   # hides the R code in the final report but still runs the code in the background.)
  bibliography: references.bib
---

# Abstract

This report investigates three research questions regarding the relationship between height, weight, gender, and physical activity in a dataset of 1,000 individuals aged 26–45. A linear regression model revealed a strong positive linear relationship between height and weight (p \< 0.001). An independent-samples t-test indicated a significant difference in average height between males and females (p \< 0.001). However, a chi-square test of independence showed no significant association between gender and physical activity (p \> 0.05). These findings provide statistical evidence about anthropometric differences across gender and highlight that physical activity levels may not vary significantly by gender in this sample.

# Introduction

This study uses data from 1,000 individuals aged 26–45, including height, weight, gender, and physical activity. To ensure reproducibility and test parameterisation, both the real dataset and a mock dataset with the same structure were used. We address three questions: whether height and weight are linearly related, whether males and females differ in average height, and whether gender is associated with physical activity. Linear regression, a t-test, and a chi-square test were applied.

# Methodology

We applied:

**Linear regression** for height–weight relationship.
$$
Model: Weight = β0 + β1Height+ϵi
$$

Linear regression analysis is a fundamental statistical method [@montgomery2021introduction].

Model: Weight = β0 + β1Height+ϵi

Test statistic: t = β1/SE(B1), following a t- distribution.

Assumptions: linearity, independence, homoscedasticity, normally distributed residuals.

**Independent samples t-test** for comparing male and female heights.The t-test and Chi-

The t-test are based on classical statistical principles developed by Fisher [@fisher1925statistical].
$$
t = \frac{\bar{X_1} - \bar{X_2}}{S_p \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}}
$$

H0: μMale = μFemale

H1:μMale umequal μFemale​

Test statistic: pooled-variance t-test.

Assumptions: normality in each group, equal variances, independence.

**Chi-square test of independence** for gender vs physical activity ndefined.

Chi-square test are based on classical statistical principles developed by Fisher [@fisher1925statistical].
$$
\chi^2 = \sum \frac{(O - E)^2}{E}
$$  

H0​: Gender and physical activity are independent.

H1: Gender and physical activity are associated.

Assumptions: expected cell counts ≥ 5.

Moreover, The analysis and data wrangling were performed in R [@R-base] using the dplyr [@R-dplyr] and ggplot2 [@R-ggplot2] packages.

Question 1 : Is there a linear relationship between height and weight?

H0 : There is no linear relationship between height and weight.

H1 : There is a linear relationship between height and weight.

```{r}
library(dplyr)
# Load dataset
df <- read.csv(params$datafile)

# 1) Linear regression: Height → Weight

# Build a regression model, coefficients table

#use lm() to build linear model
#dependent variable = weight
#independent variable = heigh
model <- lm(weight ~ height, data=df)

# Extract p-value for height
#The object model now contains the regression equation that predicts weight from height
#summary(model):This produces a detailed summary of the regression model, including a coefficients table
#$coefficients:This extracts the coefficients table from the summary output.
#[2,4]:Row 2 is corresponds to height, Column 4 is corresponds to Pr(>|t|) (the p-value)
sum_wh <- summary(model)$coefficients[2,4]


#plot() → draws a scatter plot of height (x-axis) vs. weight (y-axis).
#xlab, ylab, main → set the labels for the axes and the title of the graph.
#xlab, ylab, main → set the labels for the axes and the title of the graph.
#abline(model, col="red", lwd=2) → adds the regression line from the model (red color, line width = 2) onto the graph.
plot(df$height, df$weight,
     xlab="Height (cm)", ylab="Weight (kg)",
     main="Height vs Weight with Regression Line")
abline(model, col="red", lwd=2)


# Decision + Conclusion
if(sum_wh < 0.05){
  sum_wh_conclusion <- "There is a statistically significant positive linear relationship between height and weight."
} else {
  sum_wh_conclusion <- "There is no significant linear relationship between height and weight."
}

print(sum_wh_conclusion)

```

Interpretation

There is a statistically significant positive linear relationship between height and weight.lts

Result for question 1: Height–Weight

sum_wh = 7.57e-52 (same value, since both come from the same model)

sum_wh_conclusion = "There is a statistically significant positive linear relationship between height and weight."

p-value = 7.57e-52 :This value is extremely small (much less than 0.05).

Therefore:

Reject H₀ (null hypothesis).

Conclusion: Height and weight have a statistically significant positive linear relationship.

Question 2 : Is the average height the same for males and females? (You can assume equal variances between male and female heights.)

H₀ (null hypothesis): The average height of males and females is the same.

H₁ (alternative hypothesis): The average height of males and females is different.

```{r}
library(dplyr)

# ----------------------------------------------------------
# Load dataset (parameterised)
# ----------------------------------------------------------
df <- read.csv(params$datafile)

# ----------------------------------------------------------
# 1️ Run the independent-samples t-test (equal variances)
# ----------------------------------------------------------
t_result <- t.test(height ~ gender, data = df, var.equal = TRUE)

# Extract key values
t_p       <- t_result$p.value
t_value   <- t_result$statistic
df_value  <- t_result$parameter

# ----------------------------------------------------------
# 2️  Compute descriptive statistics by gender
# ----------------------------------------------------------
stats <- df %>%
  group_by(gender) %>%
  summarise(
    mean_height = mean(height, na.rm = TRUE),
    sd_height   = sd(height,  na.rm = TRUE),
    n           = n(),
    .groups     = "drop"
  )

male_mean   <- stats %>% filter(gender == "Male")   %>% pull(mean_height)
male_sd     <- stats %>% filter(gender == "Male")   %>% pull(sd_height)
female_mean <- stats %>% filter(gender == "Female") %>% pull(mean_height)
female_sd   <- stats %>% filter(gender == "Female") %>% pull(sd_height)

# ----------------------------------------------------------
# 3️ Generate automatic conclusion
# ----------------------------------------------------------
if (t_p < 0.05) {
  t_conclusion <- paste0(
    "Males (M = ", round(male_mean, 1),
    ", SD = ", round(male_sd, 1),
    ") were significantly taller than females (M = ",
    round(female_mean, 1),
    ", SD = ", round(female_sd, 1),
    "), t(", round(df_value, 0), ") = ", round(t_value, 2),
    ", p < 0.05."
  )
  decision <- "Reject H₀."
} else {
  t_conclusion <- paste0(
    "There was no significant difference in mean height between males and females, ",
    "t(", round(df_value, 0), ") = ", round(t_value, 2),
    ", p = ", signif(t_p, 3), "."
  )
  decision <- "Fail to reject H₀."
}

# ----------------------------------------------------------
# 4️  Display everything neatly using cat()
# ----------------------------------------------------------
cat("------------------------------------------------------------\n")
cat("Independent-Samples t-Test: Male vs Female Heights\n")
cat("------------------------------------------------------------\n\n")

cat("Hypotheses:\n")
cat("  H₀: μ(Male) = μ(Female)\n")
cat("  H₁: μ(Male) ≠ μ(Female)\n\n")

cat("Descriptive Statistics:\n")
cat("  • Male   → Mean =", round(male_mean, 1),
    "SD =", round(male_sd, 1), "\n")
cat("  • Female → Mean =", round(female_mean, 1),
    "SD =", round(female_sd, 1), "\n\n")

cat("Test Results:\n")
cat("  • t(", round(df_value, 0), ") =", round(t_value, 2), "\n")
cat("  • p-value =", signif(t_p, 3), "\n\n")

cat("Decision (α = 0.05): ", decision, "\n\n")
cat("Conclusion:\n ", t_conclusion, "\n")
cat("------------------------------------------------------------\n")

```

Results for question 2: Height by Gender

Male height: M = 178, SD = 4

Female height: M = 168, SD = 4

Test statistic: t(998) = -40.64

p-value = 7.98 × 10⁻²¹⁴ (≈ 0, much smaller than 0.05)

Therefore:\

Reject H₀ (null hypothesis).

Conclusion

There is a statistically significant difference in average height between males and females. On average, males (M = 178, SD = 4) were significantly taller than females (M = 168, SD = 4). This finding supports the biological expectation that men tend to be taller than women, and the evidence is very strong given the large sample size (N = 1000) and extremely small p-value.

Question 3 : Is there any association between gender and the amount of physical activity?

H0: There is no association between gender and physical activity.

H1: There is an association between gender and physical activity.

```{r}
library(dplyr)
library(ggplot2)

# Load dataset
df <- read.csv(params$datafile)

# Contingency table
tbl <- table(df$gender, df$phys)

# Chi-square test
chi <- chisq.test(tbl)

# Extract values
chi_p <- chi$p.value
chi_value <- chi$statistic
chi_df <- chi$parameter

# Bar plot (grouped by gender and activity level)
ggplot(df, aes(x = gender, fill = phys)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Gender vs Physical Activity",
    x = "Gender",
    y = "Count",
    fill = "Physical Activity Level"
  ) +
  theme_minimal(base_size = 14)

# Auto conclusion
if(chi_p < 0.05){
  chi_conclusion <- paste0(
    "There is a significant association between gender and physical activity, ",
    "χ²(", chi_df, ") = ", round(chi_value,2),
    ", p = ", signif(chi_p,3), "."
  )
} else {
  chi_conclusion <- paste0(
    "There is no significant association between gender and physical activity, ",
    "χ²(", chi_df, ") = ", round(chi_value,2),
    ", p = ", signif(chi_p,3), "."
  )
}

# Show contingency table and conclusion
tbl
chi_conclusion
```

Results

Contingency table (counts):

Female: Intense = 120, Moderate = 279, None = 128

Male: Intense = 127, Moderate = 236, None = 110

χ²(2) = 2.37

p-value = 0.306 (greater than 0.05)

Interpretation

Although there are slight differences in raw counts (e.g., a slightly higher proportion of females reporting moderate activity), the chi-square test indicates that these differences are not statistically significant. The observed distribution of physical activity across genders could have occurred by chance.

Therefore:

Fail to reject H₀.

Conclusion:

There is no significant association between gender and the amount of physical activity, χ²(2) = 2.37, p = 0.306 (\> 0.05). In this sample of 1,000 individuals, males and females showed similar patterns of physical activity levels, with slightly more females reporting moderate activity, but the difference was not statistically meaningful.

# Conclusion

Height–Weight Relationship

A significant positive linear relationship was found between height and weight (p = 7.57 × 10⁻⁵²), indicating that taller individuals tend to weigh more.

Average Height by Gender

Males (M = 178, SD = 4) were significantly taller than females (M = 168, SD = 4), t(998) = -40.64, p ≈ 0 (7.98 × 10⁻²¹⁴), showing a clear gender difference in height.

Gender and Physical Activity

The chi-square test showed no significant association between gender and physical activity, χ²(2) = 2.37, p = 0.306 (\> 0.05). Activity patterns were similar between males and females.

# Mock Data (for testing parameterisation)

The dataset has the same columns as file "project2025" (

ID, gender, height, weight, phys)

This guarantees your Quarto report can run without changes.

You can switch the parameter in YAML between real and mock data:

real data

params: datafile: "project2025.csv"

mock data

params: datafile: "mock_project2025.csv"

```{r}
# Set random seed for reproducibility (same results every run)
set.seed(123)

# Define sample size
n <- 1000  

# Generate mock dataset with the same structure as the real one
mock_data <- data.frame(
  ID = 1:n,
  gender = sample(c("Male", "Female"), n, replace = TRUE),         # gender gender
  height = round(rnorm(n, mean = 172, sd = 8)),                    # height  Normal(172, 8)
  weight = round(rnorm(n, mean = 70, sd = 12)),                    # weight  Normal(70, 12)
  phys   = sample(c("None", "Moderate", "Intense"), n, replace = TRUE,
                  prob = c(0.3, 0.5, 0.2))                         # phys
)

# Create a file path that works across Windows/Mac/Linux (Downloads folder)
file_path <- file.path("C:/Users/nonps/Downloads", "mock_project2025.csv")
write.csv(mock_data, file_path, row.names = FALSE)
cat("Mock data saved:", file_path, "\n")



```

Step 1: Generate the Mock Data by run code above and it will create "mock_project2025.csv"

Step 2: Switch to Mock Data in YAML

the YAML header was updated to use the mock dataset by changing params:
  datafile: "E:/StatGit/project2025.csv" to mock_project2025.csv.

```{r}
#install.packages("magick")
library(magick)
img <- image_read("E:/COMP6179-Statistic Tech/Nattapon_Pronsukwitoon_48995940/Nattapon_Pronsukwitoon_48995940/Part_1/vignettes/articles/images/Mock_data.jpg")
plot(img)

```


Step 3: : Re-render the Report

Knit/Render your quarto file again then The whole report will now run on the mock dataset. All tests (linear regression, t-test, chi-square) will still run, but the numbers will come from the fake data.

Step 4: Verify Parameterisation

If your report works with both project2025.csv (real) and mock_project2025.csv (fake). This proves that the code is parameterised correctly.\
\
these pic below is result after use mock data.

result of mock data for question 1

```{r}
library(magick)
img <- image_read("E:/COMP6179-Statistic Tech/Nattapon_Pronsukwitoon_48995940/Nattapon_Pronsukwitoon_48995940/Part_1/vignettes/articles/images/Mock_Q1.jpg")
plot(img)
```
result of mock data for question 2

```{r}
img <- image_read("E:/COMP6179-Statistic Tech/Nattapon_Pronsukwitoon_48995940/Nattapon_Pronsukwitoon_48995940/Part_1/vignettes/articles/images/Mock_Q2.jpg")
plot(img)
```


result of mock data for question 3

```{r}
img <- image_read("E:/COMP6179-Statistic Tech/Nattapon_Pronsukwitoon_48995940/Nattapon_Pronsukwitoon_48995940/Part_1/vignettes/articles/images/Mock_Q3.jpg")
plot(img)
```

